Imprecise Datapath Design: An Overclocking Approach

### 2014-01-21 TRETS2013/ReviewResponse/Response.tex

\section*{Reviewer 2}
\begin{enumerate}
  \item \Question{I believe the practical behavior is that the circuit will operate perfectly away from timing-spec corners, but that the over-clocking will simply exacerbate these. Do you believe correcting for PVT guardband by subtracting it is fair? What if arithmetic logic is more susceptible to temperature issues due to hot-spots and high toggle rate? The net effect in imprecision could be dramatically higher at process corners. I believe this should be noted/discussed (or refuted), but I think it is safe to call this future work.}

      \Response In this manuscript, for ease of discussion the PVT variation is not considered. However we agree your point that the variation is a critical issue, as it will challenge several assumptions that made in this manuscript such as the uniform carry propagation delay within a ripple-carry-adder. We plan to address this problem in the future work as presented in the Conclusion Section.

  \item \Question{A second point I would make is that you seem to have an assumption that errors do not compound. What if your arithmetic is in an IIR or used in later computations? Do the principles of the paper hold when you get an imprecise data point, then multiply or divide it in some future computation?  I think that looking at a single adder in isolation is a bit optimistic.}

      \Response In the analysis, we consider the ripple carry adder and the constant coefficient multiplier individually to reason their probabilistic behavior under overclocking and truncation. However we combine both operators into several benchmark circuits (including IIR) in our experiments. In this case, errors do propagate from different operators. However as shown in Fig.10, for instance, the overall error reduction in an FIR filter using the overclocking approach is over 2, which is predicted in Equation (17) for a single RCA. This means more benefits can be achieved when combining different arithmetic operators.

      As for utilizing the imprecise results for successive computations, we agree with the reviewers on the point that this is not included in this manuscript. We would like to investigate this in the future.

      %This motivates us to explore utilizing alternative computer arithmetic which performs computations in a most-significant-digit-first fashion.

### 2014-02-12 TRETS2013/ReviewResponse/Response.tex

\begin{document}
\maketitle

We thank the reviewers for their constructive comments and suggestions. We would like to offer the following response detailing the changes we have made as a result of their recommendations.

\section*{Reviewer 2}
\begin{enumerate}
  \item \Question{I believe the practical behavior is that the circuit will operate perfectly away from timing-spec corners, but that the over-clocking will simply exacerbate these. Do you believe correcting for PVT guardband by subtracting it is fair? What if arithmetic logic is more susceptible to temperature issues due to hot-spots and high toggle rate? The net effect in imprecision could be dramatically higher at process corners. I believe this should be noted/discussed (or refuted), but I think it is safe to call this future work.}

      \Response We would like to clarify that we run the actual (arithmetic) circuit under test on an FPGA board. So when we said in the original manuscript that we `subtract' the guard band, we meant that frequency of operation of the model is adjusted to match the observed operation. This is purely done for a fair comparison, to separate out performance boost from timing margins from error-inducing over clocking (the latter being the focus of this paper). We did not want to include the conservative timing margins within our quoted performance improvement. We have clarified this point in Section XXX in the revised manuscript.

@@

  \item \Question{A second point I would make is that you seem to have an assumption that errors do not compound. What if your arithmetic is in an IIR or used in later computations? Do the principles of the paper hold when you get an imprecise data point, then multiply or divide it in some future computation?  I think that looking at a single adder in isolation is a bit optimistic.}

      \Response This is a good point. We have not addressed this concern theoretically to date, although we do address it empirically in the paper (see Section XXX, where several chained arithmetic units are used, including an IIR filter as suggested by the reviewer; results are in line with other benchmarks). Nevertheless, we do intend to consider the theoretical limits of this behaviour, and have added this to the future work in Section XXX.

@@

  \item \Question{A third source of inaccuracy is that most FPGA timing models include crosstalk in their guardband - i.e. don't specifically do crosstalk-aware timing analysis or glitching.  Is your modeling being optimistic by looking at a circuit in isolation?  The measured/visual empirical results would incorporate this, but the model doesn't incorporate anything like it.}

      \Response Again, this is a good point. We have opted for the simplest possible model to capture the notion of over clocking, and have then reverted to experimental results to verify whether the model sufficiently captures the complexity of a real implementation. For the circuits we have used, the model prediction matches well with the empirical results, as shown in Section XXX. These benchmarks are sufficiently large to have confidence that the presence of more than one basic arithmetic core does not have a significant impact on the overall behaviour (including due to crosstalk). However, it is certainly possible that the accuracy of the model degrades as the device fills up, an issue we intend to explore in future work and have added to the list of future work in Section XXX.

@@

  \item \Question{In the introduction you claim that ``pipelining will not tend to reduce the latency of the datapath''. I don't think this is practically the case - not all paths are critical and pipelining does not imply C-slowing. For example, since a typical FPGA memory and DSP block operate grater than 500 MHz, in my preceding example you may need to pipeline some portions of the FPGA routing and logic, but you do not need to further pipeline these.}

      \Response Your point is valid if the pipeline stages are unequally distributed in time across the critical path, a non-ideal situation but - as you point out - a practical one, especially in FPGA implementation. An alternative way to improve latency in the case of your example would be to equally distribute the registers corresponding to the existing 10 ticks of latency, although we accept that this retiming is not always possible in FPGA designs. We have modified the comment to "assuming perfect pipeline stage balancing, further pipelining will not tend to reduce the latency of the datapath"; this can be found in Section XXX of the revised manuscript.

@@

  \item \Question{Wording has some comma-itis. E.g. first sentence of the intro should not have a comma in ``demonstrated that, significant performance ...''}

      \Response Thanks. This has been improved, we hope.

  \item \Question{You don't specify the speed-grade of the Xilinx device. It might be relevant if you have a down-binned device.}

      \Response The precise FPGA used was a Xilinx Virtex-6 XC6VLX240T-1FFG1156 (speed grade is -1). This has now been included in Section XXX of the revised manuscript.
\end{enumerate}

@@

\section*{Reviewer 3}
\begin{enumerate}
  \item \Question{The first sentence of the abstract is completely devoid of context, and may be the worst sentence in the entire article, given its importance. A much more appropriate sentence is the start of Section 1, Paragraph 4.}

      \Response We have modified the first two sentences of the abstract, which now read as follows:\\

      \PaperText{In this paper, we describe an alternative circuit design methodology to provide a trade-off between accuracy, performance and silicon area. We compare two different approaches that could trade accuracy for performance.}\\

@@

  \item \Question{What are the chances that an entire design with all its complexities could be accelerated in this fashion? (i.e. Could the overall clock frequency be increased?) Imprecision may be tolerable in data, but is probably not tolerable in control.}

      \Response We agree. The approach is targeted at datapath logic. In our experience, the critical path of complex signal processing systems often resides in the datapath, in which case this approach would directly improve the clock frequency. In the case where the critical path lies elsewhere, it may still be possible to improve the overall operating speed of the system-level design solely by this approach, by using multiple clock domains, which may be practical in the case of large sections of datapath with simple control structures (e.g. the filters we present) that are then interfaced to more complex control at IP boundaries. We now note this in Section XXX as an area for future work.

@@

  \item \Question{In Figure 1, you use $\mu$, whereas in the text and in Equation 4, you use $\mu_c$.  The converse is true for Section 7.4 and equations (35), (36), and (37).}

      \Response Thanks. It should be $\mu_c$ in Fig.1 and Equations (35), (36) and (37). It has been fixed.\\

  \item \Question{In Section 3.2, it would be helpful if you could illustrate or define your [-1,1) fixed-point format. Do you number your bits from k-1 to 0 or from 0 to k-1?}

 XXX -> Kan, THE TEXT BELOW IS NOT ANSWERING THE QUESTION! The question is simple: is your MSB labelled $a_0$ or $a_1$ or $a_k$ or $a_{k-1}$ - which order are the bits in??

@@

      \Response In both sections, $k$ is used to denote the word-length of the input signal, whereas $n$ is the word-length of RCA.

  \item \Question{Also in 3.3.2, the MSB is given as $2^n$, instead of $2^0$ as would be expected for the [-1, 1) fixed-point format you mentioned.}

      \Response : $2^n$ is used here because $2^i$ is used in Equation (7). Nevertheless we agree it might be confusing to put $2^n$ here, and we have changed the second paragraph of Section 3.3.2 to look as follows:\\

 XXX -> Kan, check this answer makes sense with reference to the one above about the format.

@@

      The 4-stage CSA is introduced in the fourth paragraph and Fig. 5 and Fig. 6, where we consider a variety of area constraints. Ideally with larger available area, CSA with more stages will be included in both Fig. 5 and Fig. 6. However the general trend will be similar. In previous paragraphs of this section, we only use 2-stage CSA and 3-stage CSA as examples to illustrate the design method. We have clarified this point by adding the following sentence to the fourth paragraph of Section 5.1:\\

      \PaperText{We implement CSA with all possible stage numbers within the given area specification.}\\

  \item \Question{In the experimental setup, I would be curious to know whether you used any particular constraints on the circuits-under-test to enhance their chances of success.}

      \Response  No specific constraints are applied, except the global clock constraint.

### 2014-04-25 TRETS2013/TRETS_overclocking.tex
Review 2nd round: grammer changes

% Title portion
\title{Imprecise Datapath Design: An Overclocking Approach}
\author{KAN SHI, DAVID BOLAND\\
and GEORGE A. CONSTANTINIDES
\affil{Imperial College London}
}
%\vspace{-5ex}

\begin{abstract}
%Releasing the tight accuracy requirement would potentially offer greater freedom to create a design with better performance or energy efficiency.
In this paper, we describe an alternative circuit design methodology when considering trade-offs between accuracy, performance and silicon area. We compare two different approaches that could trade accuracy for performance. One is the traditional approach where the precision used in the datapath is limited to meet a target latency. The other is a proposed new approach which simply allows the datapath to operate without timing closure. We demonstrate analytically and experimentally that on average our approach obtains either smaller errors for equivalent faster operating frequencies in comparison to the traditional approach. This is because the worst case caused by timing violations only happens rarely, while precision loss results in errors to most data. We also show that for basic arithmetic operations such as addition, applying our approach to the simple building block of ripple carry adders can achieve better accuracy or performance than using faster adder designs to achieve similar latency.
\end{abstract}

\category{B.3.3}{Performance Analysis and Design Aids}{Detail category}

@@

In order to boost the operating frequency of a datapath, the standard approaches are either to heavily pipeline the design or to reduce the precision throughout the datapath. For the former method, it should be noted that pipelining will not tend to reduce the latency of the datapath, whilst the latency in terms of clock cycles will actually increase. As a result, this method will not be applicable to many embedded applications, which often have strict latency requirements, or in any datapath containing feedback where C-slow retiming is inappropriate. For the second approach, reducing the datapath precision would reduce the latency at the cost of introducing quantization errors into the design. This has inspired research into exploiting the potential benefits of using the minimum precision necessary to satisfy a certain accuracy or performance specification~\cite{GeorgeDT11}. A brief review of existing approaches in this area will be discussed in Section~\ref{Section_Background_WL}.

@@

\begin{itemize}
	\item Detailed modeling methods of probabilistic errors generated in basic arithmetic primitives from two design scenarios to trade accuracy for performance: overclocking the datapath or truncating precisions used throughout the datapath,
	\item Exploring the accuracy-performance-area trade-offs for key arithmetic operators to illustrate the optimum architecture for a certain arithmetic function under a given requirement of silicon area,
	\item Demonstration of the proposed findings analytically and experimentally from FPGA implementations that allowing rare timing violations to happen leads to a reduction of average errors or an improvement of performance over truncating the datapath precisions to meet timing.
\end{itemize}

@@

While the aforementioned literature take advantage of the fact that only specific input patterns could cause timing errors, research on imprecise architectures take this one step further by designing simplified circuits for performance and/or energy efficiency. For example, Lu et~al. proposed a ``shrinking'' datapath that can be utilized to mimic and speculate the original logic functions~\cite{IntelSpeeding}. Kulkarni et~al. described an under-designed $2\times2$ multiplier unit, of which the worst case was replaced by a normal case based on the straight-forward Karnaugh-Map analysis~\cite{Undersigned2x2multiplier}. In both cases, reduction of area and power consumption can be achieved with the cost of accuracy. In addition, Gupta et~al. developed approximate adders at the transistor level and compared the energy efficiency of their proposed architectures over truncation of input word-length of conventional structures~\cite{Gupta2013TransCADICS}. However, we should note that using imprecise architectures means correct results would never be obtained for certain input patterns. Another limitation is that the link between the probability of output correctness and energy savings or performance improvements are rarely analyzed in current research. In addition, these techniques cannot be directly applied onto existing hardware platforms such as FPGAs.

@@

\section{Ripple Carry Adder}\label{section_RCA}
\subsection{Adder Structures in FPGAs}
Adders serve as a key building block for arithmetic operations. In general, the ripple carry adder (RCA) is the most straightforward and widely used adder structure. As such, the philosophy of our approach is first exemplified with the analysis of a RCA. We later describe how this methodology can be extended to other arithmetic operators such as CSAs and CCMs, which are commonly used in DSP applications and numerical algorithms, in Section~\ref{Section_CSA} and Section~\ref{section_CCM} respectively.

@@

 \subsubsection{Absolute Value of Overclocking Error}
For an $n$-bit RCA, let $C_{tm}$ denote the carry chain generated at bit $S_t$ with the length of $m$ bits. For a certain $f_S$, the maximum length of error-free carry propagation, $b$, is determined through (\ref{Eq:Max_CClength}). The presence of overclocking error requires $m>b$. Since the length of carry chain cannot be greater than $n$, parameters $t$ and $m$ are bounded by (\ref{Eq:t_RCA}) and (\ref{Eq:m_RCA}):
%
\begin{align}
%\small
%\footnotesize
%\scriptsize
  \label{Eq:t_RCA}  0&\leq t \leq n-b\\
    b&<m\leq n+1-t      \label{Eq:m_RCA}
%    \vspace{-1ex}
\end{align}

@@

As an example slice through the design space, in Fig.~\ref{Fig_CSA_LUT25} we record the mean value of error at outputs with respect to different operating frequencies for both design scenarios when the number of the available LUTs is set to 25. In this graph, we have labeled the optimum adder structure. Note that in this and all the following experiments within this section, in order to apply our models the input data is randomly generated following the uniform distribution.
%
\begin{figure}[t]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.3in]{./Figures/Error_LUT25.eps}
  \vspace{-2ex}
  \caption{A comparison between two design scenarios when the number of available LUT is 25. The RCA and CSA are investigated in the conventional scenario, while the RCA is explored in the proposed new scenario. The results are obtained from post place-and-route simulations on Xilinx Virtex-6 FPGAs.}
  \vspace{-2ex}
  \label{Fig_CSA_LUT25}
\end{figure}

@@

\subsection{Benchmark Circuits}
Three types of DSP designs are tested: digital filters (FIR, IIR and Butterworth), a Sobel edge detector and a direct implementation of a Discrete Cosine Transformation (DCT). The filter parameters are generated through MATLAB filter design toolbox, and they are normalized to integers for implementation. Table~\ref{Rated_Frequency_table} summarizes the operating frequency of each implemented design in Xilinx ISE14.1 when the word-length of input signal is 8-bit.
%
\begin{table}[htbp]%
\vspace{-1ex}
\tbl{Rated Frequencies of example Designs.\label{Rated_Frequency_table}}{%
\begin{tabular}{|c|c|c|}
\hline
Design & Frequency (MHz) & Description\\
\hline
FIR Filter & 126.2 & $5^{th}$ order\\
Sobel Edge Detector & 196.7 & $3\times 3$\\
IIR Filter & 140.3 & $7^{th}$ order\\
Butterworth Filter & 117.1 & $9^{th}$ order\\
DCT & 176.7 & 4-point\\
\hline
\end{tabular}}
\end{table}%

@@

Fig.~\ref{Fig:TimingMargin} also demonstrates that when the circuit is truncated, it allows the circuit to operate at a higher frequency than the frequency of full precision implementation. However, a non-uniform period change can be observed for both results. For instance, the maximum operating frequency keeps almost constant when the operand word-length reduces from 8 to 6 or from 5 to 3 in both the experimental results and those of timing analyzer. This will cause a slight deviation between our analytical model which assumes that the single bit carry propagation delay to be a constant value, as discussed in (\ref{Eq:Eo_trad_RCA}) with expression $n=b-1$. This deviation will be influenced by many factors such as how the architecture has been packed onto LUTs and CLBs. In addition, process variation might cause non-uniform interconnection delays~\cite{FPGAPV}. However, we shall see that our model remains close to the true empirical results in Section~\ref{Section_Experiments}.

@@

Table~\ref{Table:Max_Frequency} illustrates the frequency speedups for each design when the specified error tolerance varies from 0.05\% to 50\%. For all designs, we see that the overclocking scenario still outperforms the traditional scenario for each $MRE$ budget in terms of operating frequency. Likewise, the frequency speedup is higher for real image inputs than uniform inputs. The geometric mean of frequency speedups of $3.1\%$ to $21.8\%$ can be achieved by using uniform data, while $5.3\%$ to $27.6\%$ when using real image data.
%
\begin{table*}[tbp]
\vspace{-1ex}
 %% increase table row spacing, adjust to taste
   \renewcommand{\arraystretch}{1.05}
   \setlength{\tabcolsep}{3.5pt}
   \tbl{Frequency Speedups ($\%$) in Overclocking Scenario Under Various Error Budgets for Two Types of Input Data: Uniform Data (Uni) and Real Image Data from Lena (Lena).}{
%   \vspace{-2ex}
   \label{Exp_Real_Inputs}
   \centering
   %\footnotesize
   \scriptsize
   \begin{tabular}{|c|cc|cc|cc|cc|cc||cc|}
     \hline
     \multirow{2}*{\begin{tabular}{c}\textbf{Error}\\\textbf{Budget($\%$)}\end{tabular}} &
     \multicolumn{2}{c|}{\textbf{FIR}} &
     \multicolumn{2}{c|}{\textbf{Sobel}} &
     \multicolumn{2}{c|}{\textbf{IIR}} &
     \multicolumn{2}{c|}{\textbf{Butterworth}} &
     \multicolumn{2}{c||}{\textbf{DCT4}} &
     \multicolumn{2}{c|}{\textbf{Geo.Mean}} \\
     %\multirow{2}*{\begin{tabular}{c}\textbf{Geo.}\\\textbf{Mean}\end{tabular}}\\
     & {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }
     & {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }\\
     \hline
     0.05 & 4.76 &	21.43&	 6.82&	6.82&	0.95&	 1.26&	12.40&	 24.03&	 0.72&	 0.96  &3.07 & 5.32\\
     0.5  & 19.05&	21.43&	 13.64&	6.82&	0.63&	 10.06&	24.03&	 24.03&	0.48&	 12.44   &4.52 &13.45\\
     1    & 19.05&	28.57&	 13.64&	18.18&	 10.06&	16.35&	 24.03&	 24.03&	 0.48&	 12.44    &7.86 &19.10\\
     5    & 19.15&	25.53&	 18.18&	18.18&	 0.54&	0.82&	 0.63&	 0.94&	 7.66&	 12.44    &3.91    &5.36\\
     10   & 19.15&	25.53&	 10.64&	10.64&	 0.54&	1.09&	 6.92&	 13.21&	 4.91&	 4.911  &5.19  &7.19\\
     20   & 19.15&	25.53&	 5.77&	15.39&	 8.70&	8.70&	 3.26&	 3.26&	 6.70&	 10.88    &7.32    &10.39\\
     50   & 15.69&	15.69&	 10.53&	19.30&	 42.50&	50.00&	 46.74&	 54.89&	 15.06&	 19.25    &21.8 &27.59\\
     \hline
   \end{tabular}}
   \normalsize
   \label{Table:Max_Frequency}
   \vspace{-1ex}
 \end{table*}

\subsection{Area Overhead of Our Approach}
For both aforementioned design goals, it should be noted that an extra benefit of using the traditional approach is that reducing datapath precision would potentially lead to a smaller design, in comparison to our proposed overclocking scenario. Consequently, the area overheads of our approach are summarized in Table~\ref{Table:AreaFrequency} and Table~\ref{Table:AreaError} for both design goals. In our experiments, the area overhead is evaluated in terms of the number of LUTs and Slices in the FPGA technology, while similar metrics can also be employed if applying our approach into other hardware platforms. It can be seen that for a given frequency requirement with minimum possible errors, the geometric mean of area overhead ranges from $1.16\times$ and $1.15\times$ to $5.57\times$ and $4.46\times$ for LUTs and Slices, respectively. This result illustrates that although using traditional approach would lead to smaller design, the area benefits are less attractive than the performance or accuracy benefits brought by utilizing the proposed new approach, especially for not stringent frequency requirements.

An even smaller area overhead can be found for a specified error budget with the maximum clock frequencies. As seen in Table~\ref{Table:AreaFrequency}, the geometric mean of area overhead in terms of LUTs and Slices is $1.00\times\sim4.67\times$ and $1.00\times\sim3.70\times$ respectively. Notice that an area overhead of $1\times$ means both the overclocking scenario and the traditional scenario use the original precision, however the former achieves frequency speedups as seen in Table~\ref{Table:AreaError}, because a specific error budget can be tolerated.

@@

   \label{Table_AreaFrequency}

@@

   \label{Table_AreaError}

@@

While this research has shown a promising direction for sustaining performance scaling, in the future we wish to apply this approach onto arithmetic units based on alternative number representations, and explore the probabilistic behavior of these operators under a variety of process, voltage and temperature variations.

>>>>

% Title portion
\title{Imprecise Datapath Design: An Overclocking Approach}
\author{KAN SHI, DAVID BOLAND\\
and GEORGE A. CONSTANTINIDES
\affil{Imperial College London}
}
\vspace{-5ex}

\begin{abstract}
%Releasing the tight accuracy requirement would potentially offer greater freedom to create a design with better performance or energy efficiency.
In this paper, we describe an alternative circuit design methodology when considering trade-offs between accuracy, performance and silicon area. We compare two different approaches that could trade accuracy for performance. One is the traditional approach where the precision used in the datapath is limited to meet a target latency. The other is a proposed new approach which simply allows the datapath to operate without timing closure. We demonstrate analytically and experimentally that on average our approach obtains either smaller errors or equivalent faster operating frequencies in comparison to the traditional approach. This is because the worst case caused by timing violations only happens rarely, while precision loss results in errors to most data. We also show that for basic arithmetic operations such as addition, applying our approach to the simple building block of ripple carry adders can achieve better accuracy or performance than using faster adder designs to achieve similar latency.
\end{abstract}

\category{B.8.2}{Performance and Reliability}{Performance Analysis and Design Aids}

@@

In order to boost the operating frequency of a datapath, the standard approaches are either to heavily pipeline the design or to reduce the precision throughout the datapath. For the former method, it should be noted that assuming perfect pipeline stage balancing, further pipelining will not tend to reduce the latency of the datapath. As a result, this method will not be applicable to many embedded applications, which often have strict latency requirements, or in any datapath containing feedback where C-slow retiming is inappropriate. For the second approach, reducing the datapath precision would reduce the latency at the cost of introducing quantization errors into the design. This has inspired research into exploiting the potential benefits of using the minimum precision necessary to satisfy a certain accuracy or performance specification~\cite{GeorgeDT11}. A brief review of existing approaches in this area will be discussed in Section~\ref{Section_Background_WL}.

@@

\vspace{-.5ex}
\begin{itemize}
\item Detailed modeling methods of probabilistic errors generated in basic arithmetic primitives from two design scenarios to trade accuracy for performance: overclocking the datapath or truncating precisions used throughout the datapath,
\item Exploring the accuracy-performance-area trade-offs for key arithmetic operators to illustrate the optimum architecture for a certain arithmetic function under a given requirement of silicon area,
\item Demonstration of the proposed findings analytically and experimentally from FPGA implementations that allowing rare timing violations to happen leads to a reduction of average errors or an improvement of performance over truncating the datapath precisions to meet timing.
\end{itemize}
%\vspace{-2ex}

@@

While the aforementioned literature take advantage of the fact that only specific input patterns could cause timing errors, research on imprecise architectures take this one step further by designing simplified circuits for performance and/or energy efficiency. For example, Lu et~al. proposed a ``shrinking'' datapath that can be utilized to mimic and speculate the original logic functions~\cite{IntelSpeeding}. Kulkarni et~al. described an under-designed $2\times2$ multiplier unit, of which the worst case was replaced by a normal case based on the straightforward Karnaugh-Map analysis~\cite{Undersigned2x2multiplier}. In both cases, reduction of area and power consumption can be achieved at the cost of accuracy. In addition, Gupta et~al. developed approximate adders at the transistor level and compared the energy efficiency of their proposed architectures over truncation of input word-length of conventional structures~\cite{Gupta2013TransCADICS}. However, we should note that using imprecise architectures means correct results would never be obtained for certain input patterns. Another limitation is that the link between the probability of output correctness and energy savings or performance improvements are rarely analyzed in current research. In addition, these techniques cannot be directly applied onto existing hardware platforms such as FPGAs.

@@

\section{Ripple Carry Adder}\label{section_RCA}
\subsection{Adder Structures in FPGAs}
Adders serve as a key building block for arithmetic operations. In general, the ripple carry adder (RCA) is the most straightforward and widely used adder structure. As such, the philosophy of our approach is first exemplified with the analysis of an RCA. We later describe how this methodology can be extended to other arithmetic operators such as CSAs and CCMs, which are commonly used in DSP applications and numerical algorithms, in Section~\ref{Section_CSA} and Section~\ref{section_CCM} respectively.

@@

 \subsubsection{Absolute Value of Overclocking Error}
For an $n$-bit RCA, let $C_{tm}$ denote the carry chain generated at bit $S_t$ with the length of $m$ bits. For a certain $f_S$, the maximum length of error-free carry propagation, $b$, is determined through (\ref{Eq:Max_CClength}). The presence of overclocking error requires $m>b$. Since the length of the carry chain cannot be greater than $n$, parameters $t$ and $m$ are bounded by (\ref{Eq:t_RCA}) and (\ref{Eq:m_RCA}):
%
\begin{align}
%\small
%\footnotesize
%\scriptsize
  \label{Eq:t_RCA}  0&\leq t \leq n-b\\
    b&<m\leq n+1-t      \label{Eq:m_RCA}
%    \vspace{-1ex}
\end{align}

@@

As an example slice through the design space, in Fig.~\ref{Fig_CSA_LUT25} we record the mean value of error at outputs with respect to different operating frequencies for both design scenarios when the number of the available LUTs is set to 25. In this graph, we have labeled the optimum adder structure. Note that in this and all the following experiments within this section, in order to apply our models, the input data is randomly generated following the uniform distribution.
%
\begin{figure}[t]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.3in]{./Figures/Error_LUT25.eps}
  \vspace{-2ex}
  \caption{A comparison between two design scenarios when the number of available LUT is 25. The RCA and CSA are investigated in the conventional scenario, while the RCA is explored in the proposed new scenario. The results are obtained from post place-and-route simulations on Xilinx Virtex-6 FPGAs.}
  \vspace{-2ex}
  \label{Fig_CSA_LUT25}
\end{figure}

@@

\subsection{Benchmark Circuits}
Three types of DSP designs are tested: digital filters (FIR, IIR and Butterworth), a Sobel edge detector and a direct implementation of a Discrete Cosine Transformation (DCT). The filter parameters are generated through MATLAB filter design toolbox, and they are normalized to integers for implementation. Table~\ref{Rated_Frequency_table} summarizes the operating frequency of each implemented design in Xilinx ISE14.1 when the word-length of the input signal is 8-bit.
%
\begin{table}[htbp]%
\vspace{-1ex}
\tbl{Rated Frequencies of example Designs.\label{Rated_Frequency_table}}{%
\begin{tabular}{|c|c|c|}
\hline
Design & Frequency (MHz) & Description\\
\hline
FIR Filter & 126.2 & $5^{th}$ order\\
Sobel Edge Detector & 196.7 & $3\times 3$\\
IIR Filter & 140.3 & $7^{th}$ order\\
Butterworth Filter & 117.1 & $9^{th}$ order\\
DCT & 176.7 & 4-point\\
\hline
\end{tabular}}
\end{table}%

@@

Fig.~\ref{Fig:TimingMargin} also demonstrates that when the circuit is truncated, it allows the circuit to operate at a higher frequency than the frequency of full precision implementation. However, a non-uniform period change can be observed for both results. For instance, the maximum operating frequency remains almost constant when the operand word-length reduces from 8 to 6 or from 5 to 3 in both the experimental results and those of timing analyzer. This will cause a slight deviation between our analytical model which assumes that the single bit carry propagation delay to be a constant value, as discussed in (\ref{Eq:Eo_trad_RCA}) with expression $n=b-1$. This deviation will be influenced by many factors such as how the architecture has been packed onto LUTs and CLBs. In addition, process variation might cause non-uniform interconnection delays~\cite{FPGAPV}. However, we shall see that our model remains close to the true empirical results in Section~\ref{Section_Experiments}.

@@

Table~\ref{Table:Max_Frequency} illustrates the frequency speedups for each design when the specified error tolerance varies from 0.05\% to 50\%. For all designs, we see that the overclocking scenario still outperforms the traditional scenario for each $MRE$ budget in terms of operating frequency. Likewise, the frequency speedup is higher for real image inputs than uniform inputs. The geometric mean of frequency speedups of $3.07\%$ to $21.8\%$ can be achieved by using uniform data, with $5.32\%$ to $27.6\%$ when using real image data.
%
\begin{table*}[tbp]
\vspace{-1ex}
 %% increase table row spacing, adjust to taste
   \renewcommand{\arraystretch}{1.05}
   \setlength{\tabcolsep}{3.5pt}
   \tbl{Frequency Speedups ($\%$) in Overclocking Scenario Under Various Error Budgets for Two Types of Input Data: Uniform Data (Uni) and Real Image Data from Lena (Lena).}{
%   \vspace{-2ex}
   \label{Exp_Real_Inputs}
   \centering
   %\footnotesize
   \scriptsize
   \begin{tabular}{|c|cc|cc|cc|cc|cc||cc|}
     \hline
     \multirow{2}*{\begin{tabular}{c}\textbf{Error}\\\textbf{Budget($\%$)}\end{tabular}} &
     \multicolumn{2}{c|}{\textbf{FIR}} &
     \multicolumn{2}{c|}{\textbf{Sobel}} &
     \multicolumn{2}{c|}{\textbf{IIR}} &
     \multicolumn{2}{c|}{\textbf{Butterworth}} &
     \multicolumn{2}{c||}{\textbf{DCT4}} &
     \multicolumn{2}{c|}{\textbf{Geo.Mean}} \\
     %\multirow{2}*{\begin{tabular}{c}\textbf{Geo.}\\\textbf{Mean}\end{tabular}}\\
     & {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }
     & {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }\\
     \hline
     0.05 & 4.76 &	21.43&	 6.82&	6.82&	0.95&	 1.26&	12.40&	 24.03&	 0.72&	 0.96  &3.07 & 5.32\\
     0.5  & 19.05&	21.43&	 13.64&	6.82&	0.63&	 10.06&	24.03&	 24.03&	0.48&	 12.44   &4.52 &13.45\\
     1    & 19.05&	28.57&	 13.64&	18.18&	 10.06&	16.35&	 24.03&	 24.03&	 0.48&	 12.44    &7.86 &19.10\\
     5    & 19.15&	25.53&	 18.18&	18.18&	 0.54&	0.82&	 0.63&	 0.94&	 7.66&	 12.44    &3.91    &5.36\\
     10   & 19.15&	25.53&	 10.64&	10.64&	 0.54&	1.09&	 6.92&	 13.21&	 4.91&	 4.911  &5.19  &7.19\\
     20   & 19.15&	25.53&	 5.77&	15.39&	 8.70&	8.70&	 3.26&	 3.26&	 6.70&	 10.88    &7.32    &10.39\\
     50   & 15.69&	15.69&	 10.53&	19.30&	 42.50&	50.00&	 46.74&	 54.89&	 15.06&	 19.25    &21.8 &27.59\\
     \hline
   \end{tabular}}
   \normalsize
   \label{Table:Max_Frequency}
   \vspace{-1ex}
 \end{table*}

\subsection{Area Overhead of Our Approach}
For both aforementioned design goals, it should be noted that an extra benefit of using the traditional approach is that reducing datapath precision would potentially lead to a smaller design, in comparison to our proposed overclocking scenario. Consequently, the area overheads of our approach are summarized in Table~\ref{Table:AreaFrequency} and Table~\ref{Table:AreaError} for both design goals. In our experiments, the area overhead is evaluated in terms of the number of LUTs and Slices in the FPGA technology, while similar metrics can also be employed if applying our approach into other hardware platforms. It can be seen that for a given frequency requirement with minimum possible errors (Table~\ref{Table:AreaFrequency}), the geometric mean of area overhead ranges from $1.16\times$ and $1.15\times$ to $5.57\times$ and $4.46\times$ for LUTs and Slices, respectively. This result illustrates that although using traditional approach would lead to smaller design, the area benefits are less attractive than the performance or accuracy benefits brought by utilizing the proposed new approach, especially for relaxed frequency requirements.

An even smaller area overhead can be found for a specified error budget with the maximum clock frequencies. As seen in Table~\ref{Table:AreaError}, the geometric mean of area overhead in terms of LUTs and Slices is $1.00\times\sim4.67\times$ and $1.00\times\sim3.70\times$ respectively. Notice that an area overhead of $1\times$ means both the overclocking scenario and the traditional scenario use the original precision, however the former achieves frequency speedups as seen in Table~\ref{Table:AreaError}, because a specific error budget can be tolerated.

@@

@@

@@

While this research has shown a promising direction for sustaining performance scaling, in the future we wish to extend the current model by considering the propagation of error through arithmetic circuits at the component level, and to apply this approach beyond datapath designs with a variety of PVT variations.

%apply this approach onto arithmetic units based on alternative number representations, and explore the probabilistic behavior of these operators under a variety of process, voltage and temperature variations.
